A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-08-21 Manuscript/PNAS-Adams-02.MLC.tex
mlc comments

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (28--30). Nonetheless, an important question that has
yet to be considered is whether such comparisons are analytically
appropriate, and whether these statistics are, or can be, converted to
effect sizes for comparative analyses across datasets. To be statistics
representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
tree are scaled proportionally to the probability that data arise from a
BM process. Given a known \(\lambda\) value used to generate random data
on a tree, we would also expect that the mean of an empirical sampling
distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
the dispersion of \(\hat{\lambda}\) would be largest at intermediate
values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
range of \(\lambda\) with respect to tree size; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). Prior work (18) seems to support some of
these conjectures, based superficially on statistical moments for a
given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
ref. (18)). However, because the ``strength of Brownian motion'' was
simulated as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\) and not as prescribed values of
\(\lambda\) (18), interpretation of these patterns is challenging.

By contrast, for Blomberg's \(\kappa\), which is positively unbounded,
we might expect that for any \(\lambda\) used to generate data,
estimates of \(\kappa\) might follow a normal distribution, with values
distributed symmetrically about the input value. This attribute seemed
less reasonable based on the simulations performed by Münkemüller et
al.~(18), which suggested that distributions were positively skewed and
that Blomberg's \(\kappa\) might not behave as a statistic that follows
a normal distribution. However, because their simulations used a
weighted combination of simulated phylogenetic signal strengths, strong
inferences are not possible (and distributional attributes were not the
intended result of their simulations). Thus, for both Pagel's
\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
across a range of \(\lambda\) used to generate data would be valuable
for adjudicating the reliability of these statistics as effect sizes.
Furthermore, the expected values of these statistics appear to vary with
tree size (18), making comparisons across studies challenging.
Therefore, transformation of these statistics into \(Z\)-scores would
allow evaluation of the efficacy of each statistic to yield effect sizes
that could be used for comparisons of the strength of phylogenetic
signal across traits and lineages.

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases, but it is negatively-biased (particularly for small tree
sizes), and is consistently less than the input \(\lambda\) value across
most of its range (Fig. 1 black line). Second, the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig.1 red line), implying that the precision
in estimating \(\lambda\) varies across the range of input values.
Additionally, standard deviations of \(\hat{\lambda}\) are negatively
associated with tree size, and for tress of 128 species or less,
\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
are not normal across its range, but become increasingly skewed at more
extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
it is also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). Taken together these results reveal that
\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supplemental Information) reveal that incorporating
\(\hat\lambda\) in PGLS anova and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, and thus
\(\lambda\) should not be used for comparing phylogenetic signal across
datasets.

@@

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (31--33). Typically, \(\theta_{obs}\) and
\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
obtained from the distribution of \(\theta\) derived from parametric
theory. However, recent advances in resampling theory (34--37) have
shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
an empirical sampling distribution of \(\theta\) obtained from
permutation procedures.

Formalizing the suggestion of Adams and Collyer (38), an effect size for
\(\kappa\) may be found as:

@@

where \(\kappa_{obs}\) is the observed phylogenetic signal, and
\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
deviation of the empirical sampling distribution of \(\kappa\) obtained
via permutation. The empirical sampling distribution of \(\kappa\) is
first transformed via Box-Cox to better adhere to the assumption of
normality.

@@

where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
side of the equation illustrates that if \(Z_\kappa\) has already been
calculated for two sampling distributions as in equation 2, the sampling
distributions have unit variance for each of the \(Z_\kappa\)
statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
obtained from a standard normal distribution. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (35, 37).

To demonstrate the utility of \(\hat{Z}_12\), we compared \(Z_{\kappa}\)
for two ecologically-relevant traits in plethodontid salamander (Fig.
4): surface area to volume ratios (SA:V) and relative body width
(\(\frac{BW}{SVL}\)) (39, 40). While both traits contained significant
phylogenetic signal, tests based on \(\hat{Z}_{12}\) revealed that the
degree of phylogenetic signal was significantly stronger in SA:V
(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
observation may be interpreted by the fact that tropical species --
which form a monophyletic group within plethodontids -- display greater
variation in SA:V, which covaries with disparity in their climatic
niches (40). Thus, greater phylogenetic signal in SA:V is to be
expected.

@@

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\(\kappa\) (12). However, because \(\kappa\) does not scale linearly
with input levels of phylogenetic signal (Fig. 2), and its variance
increases with increasing strength of phylogenetic signal (18, 20), it
should not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\(\kappa\) to \(Z_\kappa\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (31, 41, 42), where summary statistics across datasets are
converted to standardized effect sizes for subsequent ``higher order''
statistical summaries or comparisons. As such, our approach enables
evolutionary biologists to quantitatively examine the relative strength
of phylogenetic signal across a wide range of phenotypic traits, and
thus opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (43). Thus,
\(Z_\kappa\) represents the strength of phylogenetic signal on a common
and comparable scale -- measured in standard deviations -- regardless of
the initial units and original scale of the phenotypic variables under
investigation. This means that the strength of phylogenetic signal may
be compared across datasets for continuous phenotypic traits measured in
different units and scale, because those units have been standardized
through their conversion to \(Z_\kappa\). For example, our approach
could be utilized to determine whether the strength of phylogenetic
signal (say, in response to ecological differentiation) is stronger in
morphological traits (linear traits: \(mm\)), physiological traits
(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
\(\frac{\#{displays}}{second}\)). In fact, our empirical example
provided just such a comparison, as SA:V is represented in \(mm^{-1}\)
while relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
Additionally, our method is capable of comparing the strength of
phylogenetic signal in traits of different dimensionality, as estimates
of phylogenetic signal using \(\kappa\) have been generalized for
multivariate data (16). Furthermore, tests based on \(\hat{Z}_{12}\) may
be utilized for comparing the strength of phylogenetic signal among
datasets containing a different number of species, and even for
phenotypes obtained from species in different lineages, because their
phylogenetic non-independence and observed variation are taken into
account in the generation of the empirical sampling distribution via
permutation.

This study is not the first to compare \(\lambda\) and \(\kappa\) for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \(\kappa\) values are consistent with those found in
the simulations performed by Münkemüller et al.~(18), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \(\kappa\) at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(18)). Recent work by Molina-Venegas and Rodríguez (21) found that
\(\kappa\) but not \(\lambda\) tended to inflate the estimate of
phylogenetic signal, leading to moderate type I and type II biases, if
polytomic chronograms were used. Their work more thoroughly addressed
previous observations of inflated \(\kappa\) for incompletely resolved
phylogenetic trees (18, 44). An interesting question is whether an
inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
tendency of a particular tree to inflate estimates of \(\kappa\) also
inflate the values in random permutations of a test, in which case
\(Z_\kappa\) is robust to polytomies? We repeated the analyses in Figs.
1 \& 2, adjusting trees to have 50\% collapsed nodes, per the technique
of Molina-Venegas and Rodríguez (21), and found results were consistent
(Supporting Information). This confirms that any tendency of
incompletely resolved trees to inflate \(\kappa\) as a descriptive
statistic does not inflate \(Z_\kappa\) as an effect size. Furthermore,
because comparison of effect sizes in a test is a comparison of
locations of observed values in their sampling distributions, which
would shift concomitantly because of this tendency, the \(Z_{12}\) test
statistic in equation 4 appears to be robust in spite of unresolved
trees.

@@

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
individuals of 305 species, from which species means were obtained (39,
40). A time-dated molecular phylogeny for the group (45) was pruned to
match the species in the phenotypic dataset. The phylogenetic signal in
each trait was then characterized using \(\kappa\), which was converted
to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (46,
47), and routines by the authors (\textbf{to be incorporated in
\texttt{geomorph} upon manuscript acceptance}).

@@

\leavevmode\hypertarget{ref-Glass1976}{}%
31. Glass GV (1976) Primary, secondary, and meta-analysis of research.
\emph{Educational Researcher} 5:3--8.

\leavevmode\hypertarget{ref-Cohen1988}{}%
32. Cohen J (1988) \emph{Statistical power analysis for the behavioral
sciences} (Routledge).

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
33. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
H Hedges (Russell Sage Foundation), pp 231--244.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
34. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
phenotypic change for phenotypes described by high-dimensional data.
\emph{Heredity} 115:357--365.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
35. Adams DC, Collyer ML (2016) On the comparison of the strength of
morphological integration across morphometric datasets. \emph{Evolution}
70:2623--2631.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
36. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution} 9:1772--1779.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
37. Adams DC, Collyer ML (2019) Comparing the strength of modular
signal, and evaluating alternative modular hypotheses, using covariance
ratio effect sizes with morphometric data. \emph{Evolution}
73(12):2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
38. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
Evolution, and Systematics} 50:405--425.

\leavevmode\hypertarget{ref-Baken2019}{}%
39. Baken EK, Adams DC (2019) Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
40. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
desiccation‐related morphology in plethodontid salamanders as inferred
from a novel surface area to volume ratio estimation approach.
\emph{Evolution} 74:476--486.

\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
41. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
meta-analysis} (Elsevier).

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
42. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
findings in ecology and evolution. \emph{Trends in Ecology and
Evolution} 10:236--240.

\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
43. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
San Francisco). 4th Ed.

\leavevmode\hypertarget{ref-Davies2012}{}%
44. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
resolved phylogenetic trees inflate estimates of phylogenetic
conservatism. \emph{Ecology} 93(2):242--247.

\leavevmode\hypertarget{ref-Bonett2017}{}%
45. Bonett RM, Blair AL (2017) Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, USA} 114:9936--9941.

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
46. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution} 4:393--399.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
47. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
for geometric morphometric analyses. R package version 3.3.1. Available
at: \url{https://cran.r-project.org/package=geomorph}.

>>>>

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (28--30). Nonetheless, an important question that has
yet to be considered is whether such comparisons are analytically
appropriate, and whether these statistics are, or can be, converted to
effect sizes for comparative analyses across datasets. To be statistics
representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) is a random
variable that follows the distribution of a Bernoulli probability
parameter (31); i.e., branch lengths in a tree are scaled proportionally
to the probability that data arise from a BM process. Given a known
\(\lambda\) value used to generate random data on a tree, we would also
expect that the mean of an empirical sampling distribution of
\(\hat{\lambda}\) would approximately equal \(\lambda\); the dispersion
of \(\hat{\lambda}\) would be largest at intermediate values of
\(\lambda\), \(\hat{\lambda}\) would be predictable over the range of
\(\lambda\) with respect to tree size; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (31). Prior work (18) seems to support some of these conjectures,
based superficially on statistical moments for a given tree size (mean,
variance, skewness, and kurtosis; see Fig. 2 of ref. (18)). However,
because the ``strength of Brownian motion'' was simulated as a varied
weighted-average of data simulated on trees with \(\lambda=0\) and
\(\lambda=1\) and not as prescribed values of \(\lambda\) (18),
interpretation of these patterns is challenging.

By contrast, for Blomberg's \(\kappa\), which is positively unbounded,
we might expect that for any \(\lambda\) used to generate data,
estimates of \(\kappa\) might be a random variable that follows a normal
distribution, with values distributed symmetrically about the input
value (31). This attribute seemed less reasonable based on the
simulations performed by Münkemüller et al.~(18), which suggested that
distributions were positively skewed and that Blomberg's \(\kappa\)
might not behave as a statistic that follows a normal distribution.
However, because their simulations used a weighted combination of
simulated phylogenetic signal strengths, strong inferences are not
possible (and distributional attributes were not the intended result of
their simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's
\(\kappa\), evaluation of statistical moments across a range of
\(\lambda\) used to generate data would be valuable for adjudicating the
reliability of these statistics as effect sizes. Furthermore, the
expected values of these statistics appear to vary with tree size (18),
making comparisons across studies challenging. Therefore, transformation
of these statistics into \(Z\)-scores would allow evaluation of the
efficacy of each statistic to yield effect sizes that could be used for
comparisons of the strength of phylogenetic signal across traits and
lineages.

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases, but it is negatively-biased (particularly for small tree
sizes), and is consistently less than the input \(\lambda\) value across
most of its range (Fig. 1 black line). Second, the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values (Fig.1 red line), implying that the precision
in estimating \(\lambda\) varies across the range of input values.
Additionally, standard deviations of \(\hat{\lambda}\) are negatively
associated with tree size, and for tress of 128 species or less,
\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
are not normal across its range, but become increasingly skewed at more
extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
it is also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). Taken together these results reveal that
\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supplemental Information) reveal that incorporating
\(\hat\lambda\) in PGLS ANOVA and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, and thus
\(\lambda\) should not be used for comparing phylogenetic signal across
datasets.

@@

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (32--34). Typically, \(\theta_{obs}\) and
\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
obtained from the distribution of \(\theta\) derived from parametric
theory. However, recent advances in resampling theory (35--38) have
shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
an empirical sampling distribution of \(\theta\) obtained from
permutation procedures.

Formalizing the suggestion of Adams and Collyer (39), an effect size for
\(\kappa\) may be found as:

@@

where \(\kappa_{obs}\) is the observed phylogenetic signal, and
\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
deviation of the empirical sampling distribution of \(\kappa\) obtained
via permutation. The empirical sampling distribution of \(\kappa\) is
first transformed via a Box-Cox transformation to better adhere to the
assumption of normality.

@@

where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
side of the equation illustrates that if \(Z_\kappa\) has already been
calculated for two sampling distributions as in equation 2, the sampling
distributions have unit variance for each of the \(Z_\kappa\)
statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
obtained from a standard normal distribution. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (36, 38).

To demonstrate the utility of \(\hat{Z}_{12}\), we compared
\(Z_{\kappa}\) for two ecologically-relevant traits in plethodontid
salamander (Fig. 4): surface area to volume ratios (SA:V) and relative
body width (\(\frac{BW}{SVL}\)) (40, 41). While both traits contained
significant phylogenetic signal, tests based on \(\hat{Z}_{12}\)
revealed that the degree of phylogenetic signal was significantly
stronger in SA:V (\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4).
Biologically, this observation may be interpreted by the fact that
tropical species -- which form a monophyletic group within plethodontids
-- display greater variation in SA:V, which covaries with disparity in
their climatic niches (41). Thus, greater phylogenetic signal in SA:V is
to be expected.

@@

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\(\kappa\) (12). However, because \(\kappa\) does not scale linearly
with input levels of phylogenetic signal (Fig. 2), and its variance
increases with increasing strength of phylogenetic signal (18, 20), it
should not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\(\kappa\) to \(Z_\kappa\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (32, 42, 43), where summary statistics across datasets are
converted to standardized effect sizes for subsequent ``higher order''
statistical summaries or comparisons. As such, our approach enables
evolutionary biologists to quantitatively examine the relative strength
of phylogenetic signal across a wide range of phenotypic traits, and
thus opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (44). Thus,
\(Z_\kappa\) represents the strength of phylogenetic signal on a common
and comparable scale -- measured in standard deviations -- regardless of
the initial units and original scale of the phenotypic variables under
investigation. This means that the strength of phylogenetic signal may
be compared across datasets for continuous phenotypic traits measured in
different units and scale, because those units have been standardized
through their conversion to \(Z_\kappa\). For example, our approach
could be utilized to determine whether the strength of phylogenetic
signal (say, in response to ecological differentiation) is stronger in
morphological traits (linear traits: \(mm\)), physiological traits
(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
\(\frac{\#{displays}}{second}\)). In fact, our empirical example
provided just such a comparison, as SA:V is represented in \(mm^{-1}\)
while relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
Additionally, our method is capable of comparing the strength of
phylogenetic signal in traits of different dimensionality, as estimates
of phylogenetic signal using \(\kappa\) have been generalized for
multivariate data (16). Furthermore, tests based on \(\hat{Z}_{12}\) may
be utilized for comparing the strength of phylogenetic signal among
datasets containing a different number of species, and even for
phenotypes obtained from species in different lineages, because their
phylogenetic non-independence and observed variation are taken into
account in the generation of the empirical sampling distribution via
permutation.

This study is not the first to compare \(\lambda\) and \(\kappa\) for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \(\kappa\) values are consistent with those found in
the simulations performed by Münkemüller et al.~(18), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \(\kappa\) at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(18)). Recent work by Molina-Venegas and Rodríguez (21) found that
\(\kappa\) but not \(\lambda\) tended to inflate the estimate of
phylogenetic signal, leading to moderate type I and type II biases, if
polytomic chronograms were used. Their work more thoroughly addressed
previous observations of inflated \(\kappa\) for incompletely resolved
phylogenetic trees (18, 45). An interesting question is whether an
inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
tendency of a particular tree to inflate estimates of \(\kappa\) also
inflate the values in random permutations of a test, in which case
\(Z_\kappa\) is robust to polytomies? We repeated the analyses in Figs.
1 \& 2, adjusting trees to have 50\% collapsed nodes, per the technique
of Molina-Venegas and Rodríguez (21), and found results were consistent
(Supporting Information). This confirms that any tendency of
incompletely resolved trees to inflate \(\kappa\) as a descriptive
statistic does not inflate \(Z_\kappa\) as an effect size. Furthermore,
because comparison of effect sizes in a test is a comparison of
locations of observed values in their sampling distributions, which
would shift concomitantly because of this tendency, the \(Z_{12}\) test
statistic in equation 4 appears to be robust in spite of unresolved
trees.

@@

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from
individuals of 305 species, from which species means were obtained (40,
41). A time-dated molecular phylogeny for the group (46) was pruned to
match the species in the phenotypic dataset. The phylogenetic signal in
each trait was then characterized using \(\kappa\), which was converted
to its effect size (\(Z_\kappa\)) using \texttt{geomorph} 3.3.1 (47,
48), and routines by the authors (\textbf{to be incorporated in
\texttt{geomorph} upon manuscript acceptance}).

@@

\leavevmode\hypertarget{ref-Forbes2011}{}%
31. Forbes C, Evans M, Hastings N, Peacock B (2011) \emph{Statistical
distributions} (John Wiley \& Sons).

\leavevmode\hypertarget{ref-Glass1976}{}%
32. Glass GV (1976) Primary, secondary, and meta-analysis of research.
\emph{Educational Researcher} 5:3--8.

\leavevmode\hypertarget{ref-Cohen1988}{}%
33. Cohen J (1988) \emph{Statistical power analysis for the behavioral
sciences} (Routledge).

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
34. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
H Hedges (Russell Sage Foundation), pp 231--244.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
35. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
phenotypic change for phenotypes described by high-dimensional data.
\emph{Heredity} 115:357--365.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
36. Adams DC, Collyer ML (2016) On the comparison of the strength of
morphological integration across morphometric datasets. \emph{Evolution}
70:2623--2631.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
37. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution} 9:1772--1779.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
38. Adams DC, Collyer ML (2019) Comparing the strength of modular
signal, and evaluating alternative modular hypotheses, using covariance
ratio effect sizes with morphometric data. \emph{Evolution}
73(12):2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
39. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
Evolution, and Systematics} 50:405--425.

\leavevmode\hypertarget{ref-Baken2019}{}%
40. Baken EK, Adams DC (2019) Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
41. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
desiccation‐related morphology in plethodontid salamanders as inferred
from a novel surface area to volume ratio estimation approach.
\emph{Evolution} 74:476--486.

\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
42. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
meta-analysis} (Elsevier).

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
43. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
findings in ecology and evolution. \emph{Trends in Ecology and
Evolution} 10:236--240.

\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
44. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
San Francisco). 4th Ed.

\leavevmode\hypertarget{ref-Davies2012}{}%
45. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
resolved phylogenetic trees inflate estimates of phylogenetic
conservatism. \emph{Ecology} 93(2):242--247.

\leavevmode\hypertarget{ref-Bonett2017}{}%
46. Bonett RM, Blair AL (2017) Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, USA} 114:9936--9941.

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
47. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution} 4:393--399.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
48. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
for geometric morphometric analyses. R package version 3.3.1. Available
at: \url{https://cran.r-project.org/package=geomorph}.
